---
title: Effective Use of "data.table"
author: Oscar Riordan
date: '2020-11-28'
slug: effective-use-of-data-table
categories:
  - guide
  - package
  - data.table
tags: []
---

# Introduction

data.table is a complex but rewarding package to use. It is by far the most
efficient out there for data processing and manipulation. However it can be 
tricky to get to grips with integrating it in to work flows, especially with 
many packages being built to fit with tidyverse and offering examples using
that collective of packages.

It's efficient style and operation are worth the extra time to learn how to use 
it in a readable and effective way. With that being said there are certain 
things I wish I had known when starting and that I hope may be helpful to 
others.

# The Basics

```{r}
library(data.table)
library(magrittr)

dt <- setDT(copy(mtcars))
```

## The most important operator

Firstly I think it is necessary to understand the assignment operator. This can
be used thusly:
```{r}

dt[, hp_to_wt := hp / wt]

head(dt)
```

```{r}
dt[hp_to_wt < 40, hp_to_wt := 20]

head(dt)
```

It is the assignment operator and it is the most efficient  

## When and when not to use vectors

The rules for when vectors can and can't be used to select and manipulate data
is a little opaque. It can seem at first that lists (including the shothand 
version offered by data.table: ".()") and vectors are interchangable except when
they are not.





